### 2014-02-08 abstract.tex
Tyler's comments

\section*{Abstract}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second leading cause of cancer-related death.
Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
popularity to characterize PCa.  ARFI imaging has the potential to aid in PCa
diagnosis and management by evaluating the structural composition of prostate
zones and tumors based on their stiffness.  In this study, \invivo MR and ARFI
imaging datasets were compared to one another and with gross pathology
measurements made immediately post radical prostatectomy.  Images were manually
segmented in 3D Slicer to delineate the central gland (CG) and prostate
capsule, and 3D models were rendered in 3D Slicer to evaluate zonal anatomy
dimensions and volumes.  Both imaging modalities showed moderate correlations
between estimated organ volume and gross pathologic weights.  ARFI and MR total
prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but
ARFI images yielded prostate volumes that were, on average, larger
(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images,
primarily due to over-estimation of the lateral dimension of the total gland
(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), while differences
in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
\ARFImrTotalAntPostStdPct\%, anterior-to-posterior, and
\ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%, apex-to-base).
The ARFI and MR CG volumes were also well-correlated (R$^2$ =
\MRarfiVolCentralRsq).  CG volume differences were attributed to
under-estimation of the apex-to-base axis (\ARFImrCentralApexBaseMeanPct $\pm$
\ARFImrCentralApexBaseStdPct\%), and over-estimation bias in the lateral
dimension (\ARFImrCentralLatLatMeanPct $\pm$ \ARFImrCentralLatLatStdPct\%).
ARFI prostate imaging yielded prostate volumes and dimensions that were well
correlated with MR T2WI estimates, with biases in the lateral dimension due to
poor contrast caused by peri-prostatic fat.  ARFI imaging is a promising
low-cost, real-time imaging modality that can compliment mpMRI for diagnosis,
treatment planning and management of PCa.

>>>>

\section*{Abstract}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second leading cause of cancer-related death.
Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
popularity to characterize PCa.  ARFI imaging has the potential to aid in PCa
diagnosis and management by evaluating the structural composition of prostate
zones and tumors based on their stiffness.  In this study, \invivo MR and ARFI
imaging datasets were compared to one another and with gross pathology
measurements made immediately after radical prostatectomy.  Images were
manually segmented in 3D Slicer to delineate the central gland (CG) and
prostate capsule, and 3D models were rendered to evaluate zonal anatomy
dimensions and volumes.  Both imaging modalities showed good correlation
between estimated organ volume and gross pathologic weights.  Ultrasound and MR
total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq),
but ARFI images yielded prostate volumes that were larger
(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images,
primarily due to overestimation of the lateral dimension
(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), while differences
in the other dimensions were less significant (\ARFImrTotalAntPostMeanPct $\pm$
\ARFImrTotalAntPostStdPct\%, anterior-to-posterior, and
\ARFImrTotalApexBaseMeanPct $\pm$ \ARFImrTotalApexBaseStdPct\%, apex-to-base).
The ARFI and MR CG volumes were also well-correlated (R$^2$ =
\MRarfiVolCentralRsq).  CG volume differences were attributed to
ARFI underestimation of the apex-to-base axis (\ARFImrCentralApexBaseMeanPct $\pm$
\ARFImrCentralApexBaseStdPct\%), and ARFI overestimation bias in the lateral
dimension (\ARFImrCentralLatLatMeanPct $\pm$ \ARFImrCentralLatLatStdPct\%).
ARFI imaging yielded prostate volumes and dimensions that were well-correlated
with MR T2WI estimates, with biases in the lateral dimension due to poor
contrast caused by peri-prostatic fat.  ARFI imaging is a promising low-cost,
real-time imaging modality that can compliment mpMRI for diagnosis, treatment
planning and management of PCa.

### 2014-02-08 fig_mr_arfi_weight.tex
Tyler's comments

\begin{figure}[htb!]
\centering
\begin{tabular}{lll}
\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol} \\
(a) Path Weight : Path Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Path Volume \\
\end{tabular}
%\begin{tabular}{ll}
%\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol_no4} &
%\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol_no4} \\
%(d) Image Volume : Prostate Weight (-4) & (e) Image Volume : Path Volume (-4) \\
%\end{tabular}
\caption{Tri-axial pathology measurements were used to make an ellipsoidal
    prostate volume approximation based on gross pathology axis measurements,
    which was moderately well-correlated with the excised prostate weights (a,
    R$^2$ = \pathVolWeightRsq).  T2WI MR (blue, X) showed a moderate
    correlation between the reconstructed volumes and prostate weight (R$^2$ =
    \weightMRrsq), while volumes reconstructed from ARFI images (green, O)
    showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Weaker
    correlations existed between both T2WI MR and ARFI image volumes and
    approximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
    and \pathVolARFIrsq, respectively) (c).  There was one pathology specimen
    that had an erroneous prostate volume and weight recorded, which has been
    excluded from the linear regressions (magenta circles).}
\label{fig:mr_arfi_weight}
\end{figure}

>>>>

\begin{figure}[htb!]
\centering
\begin{small}
\begin{tabular}{lll}
\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol} \\
(a) Prostate Weight : Prostate Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Prostate Volume \\
\end{tabular}
\end{small}
%\begin{tabular}{ll}
%\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol_no4} &
%\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol_no4} \\
%(d) Image Volume : Prostate Weight (-4) & (e) Image Volume : Path Volume (-4) \\
%\end{tabular}
\caption{Tri-axial pathology measurements were used to make an ellipsoidal
    prostate volume approximation based on gross pathology axis measurements,
    which was moderately well-correlated with the excised prostate weights (a,
    R$^2$ = \pathVolWeightRsq).  T2WI MR (blue, X) showed a moderate
    correlation between the reconstructed volumes and prostate weight (R$^2$ =
    \weightMRrsq), while volumes reconstructed from ARFI images (green, O)
    showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Weaker
    correlations existed between both T2WI MR and ARFI image volumes and
    approximated ellipsoidal prostate pathology volumes (R$^2$ =
    \pathVolMRrsq~and \pathVolARFIrsq, respectively) (c).  There was one
    pathology specimen that had an erroneous prostate volume and weight
    recorded, which has been excluded from the linear regressions (magenta
    circles).}
\label{fig:mr_arfi_weight}
\end{figure}

### 2014-02-08 image_proc_methods.tex
Tyler's comments

\subsubsection{Ultrasound Image Segmentation and Modeling}
Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
using the \verb+draweffect+ tool to model and measure the prostate capsule and
CG volume and tri-axial dimensions (Figure~\ref{fig:arfi_segs}).  The sagittal
imaging plane was chosen for segmentation because it has a greater number of
easily identifiable anatomical structures:  seminal vesicles in the left and
right posterior base, urethra in the mid apex, surrounding fibromuscular tissue
in the anterior, and rectum in the posterior. Prostate capsule was segmented in
B-mode as this modality offers both high contrast between prostatic tissue and
surrounding peri-prostatic fat, and better SNR in the anterior capsule region
than ARFI images. ARFI images were used for CG segmentation.  Prior to
segmentation, both the B-mode and ARFI image stacks were down-sampled along the
sagittal axis by a factor of 20 (i.e., from 360 slices per image stack to 18
slices per image stack) to reduce the number of manually-segmented frames.

While segmentations were performed along the sagittal plane, label map
intersections in the axial and coronal views were used during active creation
of the sagittal segmentation to further improve segmentation accuracy.
Parameters used for 3D model rendering were identical to those used for MR
modeling (Table~\ref{tab:3dslicer}).  The modeled capsule helped guide the CG
segmentation by enabling capsule model intersections in the three orthogonal
planes during CG segmentation (Figure~\ref{fig:arfi_segs}). The capsule provided
an outer boundary for the CG segmentation. The capsule and CG models were then
measured using the same procedure as the MR images.

>>>>

\subsubsection{Ultrasound Image Segmentation and Modeling}
Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
using the \verb+draweffect+ tool to model and measure the prostate capsule and
CG volume and tri-axial dimensions (Figure~\ref{fig:arfi_segs}).  The sagittal
imaging plane was chosen for segmentation because it has a greater number of
easily identifiable anatomical structures:  seminal vesicles in the left and
right posterior base, urethra in the mid apex, surrounding fibromuscular tissue
in the anterior, and rectum in the posterior. Prostate capsule was segmented in
B-mode as this modality offers both high contrast between prostatic tissue and
surrounding peri-prostatic fat, and better SNR in the anterior capsule region
than ARFI images. ARFI images were used for CG segmentation.  Prior to
segmentation, both the B-mode and ARFI image stacks were down-sampled along the
sagittal axis to a slice thickness of 3 mm to reduce the number of
manually-segmented frames (18 vs. 360 slices / image stack) and match the MRI
slice thickness.

While segmentations were performed along the sagittal plane, label map
intersections in the axial and coronal views were used during active creation
of the sagittal segmentation to further improve segmentation accuracy.
Parameters used for 3D model rendering were identical to those used for MR
modeling (Table~\ref{tab:3dslicer}).  The modeled capsule helped guide the CG
segmentation by enabling capsule model intersections in the three orthogonal
planes during CG segmentation (Figure~\ref{fig:arfi_segs}). The capsule
provided an outer boundary for the CG segmentation. The capsule and CG models
dimensions were then measured using the same procedure as the MR images.
